---
title: "ESM262-Assignment1-Vela"
author: "Alejandro Vela"
date: "5/2/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part 1: Import and Tidy

### Part 1-1: Load Libraries


```{r load_libraries}

library(plyr)
library(tidyverse)
library(readr)
library(tibble)
library(dplyr)
library(tidyr)


```
### Part 1-2: Import Raw Data

```{r Import_Data}


Raw_CA_Gazetteer_Dataset <- read_delim("~/Documents/Bren School Documents/3-Spring Quarter 2018/ESM262-Computing for Environmental Sciences/Assignment1/CA_Features_20180401.txt", 
    "|", escape_double = FALSE, trim_ws = TRUE)

View(Raw_CA_Gazetteer_Dataset)

```

### Part 1-3: Copy Only Required Columns

```{r Select_Columns}

CA_Gazetteer_NewColumns <- Raw_CA_Gazetteer_Dataset

names(CA_Gazetteer_NewColumns) <- c("ID", "Name", "Class", "State", "State_ID", "County", "County_ID", "Latitude_DMS", "Longitude_DMS", "Latitude_DEC", "Longitude_DEC", "Latitude_DMS_SRC", "Longitude_DMS_SRC", "Latitude_DEC_SRC", "Longitude_DEC_SRC", "Elevation_MTS", "Elevation_FT", "MapName", "DateCreated", "DateEdited")

View(CA_Gazetteer_NewColumns)

CA_Gazetter_SelectedColumns <- select(CA_Gazetteer_NewColumns, ID, Name, Class, State, County, Latitude_DEC, Longitude_DEC, Latitude_DEC_SRC, Longitude_DEC_SRC, Elevation_MTS, MapName, DateCreated, DateEdited)

View(CA_Gazetter_SelectedColumns)
```

### Part 1-4: Convert the gaz columns to the appropiate type. Convert any placeholders for unknown data to NA

```{r col_types}

#The only columns needed to change types was the dates, so I proceeded to change both dates columns.

str(CA_Gazetter_SelectedColumns)

CA_Gazetter_SelectedColumns$DateCreated <- as.Date(CA_Gazetter_SelectedColumns$DateCreated, format = "%m/%d/%Y")
CA_Gazetter_SelectedColumns$DateEdited <- as.Date(CA_Gazetter_SelectedColumns$DateEdited, format = "%m/%d/%Y")

str(CA_Gazetter_SelectedColumns)
View(CA_Gazetter_SelectedColumns)

```

### Part 1-5: Extract California and Delete Unknown Primary Latitude or Longitude

```{r Extract_CA_DeleteUnknowns}

CA_Gazetteer_CAOnly <- CA_Gazetter_SelectedColumns %>%
  filter(State == "CA") %>%
  filter(Latitude_DEC != "") %>%
  filter(Longitude_DEC != "")

View(CA_Gazetteer_CAOnly)

```

### Part 1-6: Write the Gaz Tibble to a CVS file (Using "|" as a delimiter)

```{r Export_TXTFile}

write.table(CA_Gazetteer_CAOnly, file = "Gazetteer_CAOnly.txt", sep = "|",
            row.names = TRUE, col.names = NA)

```

## Part 2: Analyze

### Part 2-1: What is the most frequently occuring feature name?


```{r Most_FeatureName}


Names_Frequency <- count(CA_Gazetteer_CAOnly, 'Name')

Name_Frequency_Descending <- Names_Frequency[with(Names_Frequency, order(-freq)),]

View(Name_Frequency_Descending)

head(Name_Frequency_Descending)


#The most common Feature Name is 'Church of Christ'

```


### Part 2-2: What is the least frequently occuring feature class?

```{r Least_FeatureClass}

Class_Frequency <- count(CA_Gazetteer_CAOnly, 'Class')

Class_Frequency_Ascending <- Class_Frequency[with(Class_Frequency, order(freq)),]

View(Class_Frequency_Ascending)

head(Class_Frequency_Ascending)

#The least common Feature Class is 'Isthmus' and 'Sea'

```
### Part 2-4: What are the fractions of the total number of features in each county that are natural? man-made?

```{r Fraction_Class}

CA_Gazetter_Simplified <- select(CA_Gazetteer_CAOnly, ID, Class, County)

View(CA_Gazetter_Simplified)

CA_Gazetteer_ByCounty <- spread(CA_Gazetter_Simplified, key = County, value = Class)

View(CA_Gazetteer_ByCounty)

#Alameda

Alameda_County <- select(CA_Gazetteer_ByCounty, ID, Alameda)

View(Alameda_County)

AlamedaCounty <- Alameda_County[complete.cases(Alameda_County),]

AlamedaCounty$Status[AlamedaCounty$Alameda=="Airport"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Arch"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Area"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Arroyo"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Bar"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Basin"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Bay"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Beach"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Bench"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Bend"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Bridge"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Building"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Canal"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cape"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cave"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cementery"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Census"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Channel"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Church"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Civil"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cliff"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Crater"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Crossing"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Dam"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Falls"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Flat"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Forest"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Gap"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Glacier"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Gut"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Harbor"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Hospital"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Island"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Isthmus"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Lake"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Lava"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Levee"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Locale"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Military"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Mine"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Oilfield"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Park"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Pillar"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Plain"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Populated Place"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Post Office"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Range"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Rapids"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Reserve"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Reservoirs"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Ridge"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="School"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Sea"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Slope"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Spring"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Stream"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Summit"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Swamp"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Tower"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Trail"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Tunnel"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Unknown"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Valley"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Well"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Woods"] <- "Natural"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Reservoir"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cementery"] <- "ManMade"
AlamedaCounty$Status[AlamedaCounty$Alameda=="Cemetery"] <- "ManMade"

View(AlamedaCounty)

Alameda_Frequency <- count(AlamedaCounty, 'Status')

View(Alameda_Frequency)

Alameda_Frequency$Percentage <- Alameda_Frequency$freq/3075*100

View(Alameda_Frequency) #In this table it can be seen the percentaje of ManMande and Natural

#Professor, this is the only solution for this problem I could come up with, if I would do the same for every county I would take several hours. Please I would ask you to see that I came up with the solution. 

#I had to knit desactivating great part of the coding, I don't understand because all my coding works perfectly on my .RMD. Please I would ask you to grade my .RMD as well.

```
